### 2017-09-06 paper/rebuttal.txt
paper updates

Rebuttal and comments on reviewer comments:

Thank you to the reviewers for there reviews and very useful comments.

Please find in the following a short log on how we incorporated the comments in our paper, and reasons 
why we did not follow some of the reviewer's proposals.

@Reviewer #1:

Unfortunately, the paper is already scratching the 21 page limit, so that we were hardly able to add more detail.

We added a short explanation of reflexive transitive closure of LTSs, and explained Isabelle's non-standard set comprehension syntax.
We also added a more detailed explanation of the Hoare-triple and separation logic syntax used for the correctness theorem.

We changed the wording for the presentation of Isar, to not present it as unusual. 
We also added the argument of the community size behind MLton vs. gcc.

Chargueraud et al. actually used their methods to verify graph algorithms, while Chlipala et al. 
only applied refinement in a quite specific setting to data-structure refinement for databases. 
I'm not aware of any other application of Fiat so far. As this paper is mainly about human-readable formalizations of (graph) algorithms,
and not so much about automated refinement techniques, we did not include Fiat into the related work discussion.

@Reviewer2:

We tried to clarify in the introduction that we only formalize a O(V^2E) bound. 
We decided not to integrate this in the abstract, in order to keep it short.
Also, in the related work section, we added some discussion on our method of estimating complexity, 
compared to other methods of Nipkow and Chargueraud.

We clarified that the transformation to satisfy our constraints is linear time, and not part of our formalization.

We explained the non-standard set-comprehension syntax.

We re-phrased the paragraph mentioning that we need to provide more powerful combinators in the Refinement Framework.

We gave a pointer to our ITP'12 paper \sec 2.3, where a more detailed explanation of total vs partial correctness in the Refinement Framework can be found.

Unfortunately, the page-limit does not allow a more detailed explanation of the induction proof for the Edmonds-Karp complexity.

We slightly changed Sec 5, and added a discussion about our complexity analysis compared to other methods to related work.

We tried to re-phrase the informal explanation of push-relabel algorithms to make it easier to read. We think that 
"simple path" is standard enough that we do not need to define it explicitly. At least from the context it should be 
quite clear (now) that it is a path with no repeated vertexes.

We added a footnote to show how much we loosened the constant factors in our analysis, compared to Cormen.

We added some comments on our matrix implementation of graphs, and its deficits on sparse graphs.

We added some explanation of the syntax used in the main correctness statement (Hoare triple, sep-logic, ...)

We added some missing curves to the benchmark charts, and explained why some are still missing.

We added Chen and Levys work to related work, we also discuss Chargueraud and Pottier for their complexity analysis techniques.
However, we did not want to discuss any formalization of something like a graph algorithm, to keep the related work section concise 
(and within the page limit).

@Reviewer 3:

We did not want to over-emphasize these rather low-level and easy to fix parser and overflow errors in the reference 
implementations, thus we kept it as footnote.

We explained our Hoare-triple syntax. I, personally, am not a fan of too overloaded syntactic sugar, and there is already a lot of that in the Refinement Framework.